---
title: 'Problem Set #1'
author: "Erica Criollo"
date: "2025-01-29"
output: pdf_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.align = 'center',
  message = FALSE,
  warning = FALSE
)

# Loading packages and importing data

library(tidyverse)
library(knitr)
library(kableExtra)
library(haven)
library(dplyr)
library(tableone)


df <- read_sav("UGA_R9.data_.final_.wtd_release.19Jun22_Updated.7Nov23 (1).sav")
#View(df)

```


## Introduction
This report will be utilizing data from the Round 9 of the Afrobarometer survey in Uganda. This survey was conducted on June 19, 2022 and had 2400 respondents. 


```{r}
# Creating a new data frame with variables including urban/rural, Ages, 
#Gender, Language, Questions on China, Questions on US

PSData <- subset(df, select=c(RESPNO, URBRUR, Q1, Q100, Q2, Q78A, Q78B))

```


## Descriptive Statistics
From the survey, 50% were men and 50% were women. Of all respondents, 43% ranged from 18-30 years old, 38% ranged from 30-50 years old, 16% ranged from 50-70 years old, and 3% were older than 70 years old. The majority of respondents lived in rural areas at 73% while the rest, 27%, lived in urban areas. When analyzing language, the three most popular languages spoken among respondents are Luganda (25%), Lusoga (12%), and Runyankore (10%). 



```{r}
#Exploring new data frame and summarizing statistics

#str(PSData)
#summary(PSData)

#Finding Percent of Gender (1=Man 2 =Woman )
Gender <- table(PSData$Q100)
Gender <-prop.table(Gender)*100
#View (Gender)

#Change Age from numerical to categorical

#summary(PSData$Q1)
PSData$Q1 <- cut(PSData$Q1,
                 breaks = c(18, 30, 50, 70, 85),
                 labels = c("18-30", "30-50", "50-70", "More than 70"))

#Finding Percent of Ages

Age <- table(PSData$Q1)
Age <-prop.table(Age)*100
#View (Age)

# Finding Percent of Urban/Rural (1=Urban 2=Rural)

Urbrul <- table(PSData$URBRUR)
Urbrul <-prop.table(Urbrul)*100
#View (Urbrul)

#Finding Percent of Language

Language <- table(PSData$Q2)
Language <-prop.table(Language)*100
#View (Language)


```




## Attitudes


This report uses the variables Q78A and Q78B which measures participants response to the question: "Do you think that the economic and political influence of [China or the U.S] on Uganda is mostly positive, mostly negative, or haven’t you heard enough to say?", to find differences in opinion. 


### Economic and Political Influence of China


Most of the participants view on China’s political and economic influence on Uganda is positive, with 47% expressing favorable opinions along Somewhat Positive (27%) and Very Positive (20%). Yet, 26% of respondents responded as being uncertain and negative perceptions are an overall 24%, with Somewhat Negative (12%) and Very Negative (12%) responses.


```{r tab 1}


#Exploring Q78A Variable

#str(PSData$Q78A)
#summary(PSData$Q78A)


#Creating table of relative frequency

AttChina <- table(PSData$Q78A)/length(PSData$Q78A)*100
#str(AttChina)
AttChina <- as.data.frame(AttChina) 
#View(AttChina)
#summary(AttChina)


#Changing Labels and Aesthetics

colnames(AttChina) <- c("Attitudes", "Percent")
AttChina <- AttChina %>%
  mutate(Attitudes = case_when(
    Attitudes == 1 ~ "Very Negative",
    Attitudes == 2 ~ "Somewhat Negative",
    Attitudes == 3 ~ "Neither Positive Nor Negative",
    Attitudes == 4 ~ "Somewhat Positive",
    Attitudes == 5 ~ "Very Positive",
    Attitudes == 8 ~ "Refused to Answer",
    Attitudes == 9 ~ "Don't know",
    TRUE ~ as.character(Attitudes)
  ))


AttChina %>%
  kable(
    caption = 'Attitudes of Economic and Political Influence of China (%)',
    digits = 1L,
    format = 'latex',
    booktabs = TRUE,
    
  )


```



### Economic and Political Influence of the United States of America


Most of the participants view on the United State’s political and economic influence on Uganda is uncertain with 34% of all respondents answering "don't know". For positive attitudes, Somewhat Positive (28%) and Very Positive (17%) make 45% of total positive feelings. Negative perceptions are an overall 15%, with Somewhat Negative (10%) and Very Negative (5%) responses.



```{r tab2}

#Exploring Q78B Variable

#str(PSData$Q78B)
#summary(PSData$Q78B)


#Creating table of relative frequency

AttUSA <- table(PSData$Q78B)/length(PSData$Q78B)*100
#str(AttUSA)
AttUSA <- as.data.frame(AttUSA) 
#View(AttUSA)
#summary(AttUSA)


#Changing Labels and Aesthetics

colnames(AttUSA) <- c("Attitudes", "Percent")
AttUSA <- AttUSA %>%
  mutate(Attitudes = case_when(
    Attitudes == 1 ~ "Very Negative",
    Attitudes == 2 ~ "Somewhat Negative",
    Attitudes == 3 ~ "Neither Positive Nor Negative",
    Attitudes == 4 ~ "Somewhat Positive",
    Attitudes == 5 ~ "Very Positive",
    Attitudes == 8 ~ "Refused to Answer",
    Attitudes == 9 ~ "Don't know",
    TRUE ~ as.character(Attitudes)
  ))


AttUSA %>%
  kable(
    caption = 'Attitudes of Economic and Political Influence of the United States of America (%)',
    digits = 1L,
    format = 'latex',
    booktabs = TRUE,
    
  )



```


### T-Test


In order to find the difference of means between attitudes toward influence of China and the United States on Uganda, a t-test was calculated. Assuming 5% significance for a two-tailed t-test, the analysis rejects the null that there is no difference between attitudes being that the p-value (1.661e-07) is less than 0.05. With t = -5.258, the analysis concludes that there is a statistically significant difference between attitudes. The mean difference (-0.2029) indicates that on average, respondents felt more negative about China's influence than U.S influence on Uganda.


```{r}

PSData = 
  PSData %>%
  mutate(
    across(
      Q78A:Q78B,
      ~ if_else(.x %in% 1:5, .x, NA)
    )
  )

t.test(PSData$Q78A, PSData$Q78B, paired = TRUE)

```


